---
title: "Task 1 - PCA Analysis of Global Climate Data"
author: "Logan Ossentjuk"
output: 
  html_document:
    code_folding: hide
---

# Overview: 
This brief examines variables affecting climate from nearly 230 countries . The data is subsetted to identify key interest variables before running a principle component analysis. The results of this analysis demonstrate relationships and correlations between chosen variables. The results are then shown in a biplot and conclusions are drawn. 

**Data Citation**: Compiled and provided by @zander_venter on Kaggle, (all data acquired via Google Earth Engine https://earthengine.google.com/)

```{r setup, include=TRUE, message= FALSE, warning=FALSE}
knitr::opts_chunk$set(echo = TRUE, message = FALSE, warning = FALSE)
library(tidyverse)
library(janitor)
library(here)
library(ggfortify)
library(patchwork)
```

```{r, include= TRUE}
### Data Wrangling and Subsetting
world_env <- read_csv(here('data', 'world_env_vars.csv')) %>% 
  clean_names() %>% 
  drop_na()

world_env_sub <- world_env %>% 
  select(country, accessibility_to_cities, cropland_cover, tree_canopy_cover, rain_mean_annual,temp_annual_range, temp_mean_annual, temp_seasonality, wind, cloudiness, elevation) %>% 
  rename(
    "Country" = country,
    "Accessibility To Cities" = accessibility_to_cities,
    "Elevation" = elevation,
    "Cropland Cover" = cropland_cover,
    "Tree Canopy Cover" = tree_canopy_cover,
    "Annual Mean Rainfall" = rain_mean_annual,
    "Annual Temperature Range" = temp_annual_range,
    "Annual Mean Temperature" = temp_mean_annual,
    "Temperature Seasonality" = temp_seasonality,
    "Wind" = wind, 
    "Cloudiness" = cloudiness)


###Creating PCA set
world_env_pca <- world_env_sub %>% 
  select(-Country) %>% 
  scale() %>% #scales variables based on ranges onto the same spectrum
  prcomp(scale = TRUE)


#world_env_pca
```

### PCA Analysis 
```{r}
autoplot(world_env_pca, 
         data = world_env_sub, 
         loadings = TRUE,
         colour = 'darkgreen',
         loadings.label = TRUE,
         loadings.colour = 'goldenrod3',
         loadings.label.colour = 'black',
         loadings.label.vjust = -0.5) + 
        #scale_color_manual(values = c('blue', 'purple', 'orange')) +
        theme_grey()
```
<br> 
**Figure 1. **  Illustrates principle components analysis results of the acquired climate data. Each country is shown by a green dot. Orange arrows demonstrate chosen variables of interest where increased arrow length denotes an increased role in explaining the variation within this data set. 


# Summary of Findings: 
Cloudiness and Wind are negatively correlated while Cloudiness, Tree Canopy Cover, and Annual Mean Rainfall are all positively correlated. It makes sense that wind and cloudiness would negatively correlated as increased wind would decrease cloudiness. It is also logical that with increased Annual Mean Rainfall, Tree Canopy Cover would increase as plants provided with ample water will grow more lush. It is interesting here as well that Cropland Cover and Accessibility to Cities are negatively correlated. This is likely due to differing land use, where cropland is often outside of cities. However, given the poor strength of accessibility to cities, in PCA 1 or 2 it is difficult to infer this relationship based on the data. 

**Other general findings: **

- Tree Canopy Cover and Accessibility to Cities are negatively correlated along with Wind. This likely points to the role of deforestation in building cities and the resultant increase in Wind resulting from a smoother land surface
<br>

- Cropland Cover and Annual Mean Temperature are negatively correlated, possibly pointing toward a middle ground temperature that cropland is most abundant. 
<br>

- Temperature Seasonality and Elevation are relatively positively correlated, pointing toward weather trends that shift as elevation increases. 

<br>

 Tree canopy was positively correlated mean and wettest month rainfall, while negatively correlated with temperature seasonality and temperature range. This aligns with our knowledge that trees need a high level of precipitation to thrive. While we cannot infer causality from the correlation, it raises an interesting questions about if the presence of forests serves as a buffer against temperature extremes or if temperature extremes undermine trees' survivability.
- Driest month of rainfall was negatively correlated with temperature range, perhaps pointing to the fact that desert-like environments have large temperature swings throughout the day.
- Rain seasonality was almost completely unrelated to temperature seasonality.
- The variation in accessibility to cities was not captured very well by PCA 1 or PCA 2, indicating a weak relationship between accessibility to cities and the other variables analyzed.

```{r}
