---
title: "Task 1 - "
author: "Logan Ossentjuk"
output: 
  html_document:
    code_folding: hide
---

You only need to **use one dataset to perform PCA, create a biplot, and interpret the results.** Whichever dataset you choose, create a finalized HTML (knitted from .Rmd) that includes: 
**A useful descriptive introductory summary (3 - 4 sentences) that helps the audience understand the data (include a citation as necessary) and what you’ll be exploring**
**All of your organized and well-annotated code (with warnings/messages suppressed) you wrote to wrangle data then run PCA, and to create a professional looking PCA biplot that appears (with a caption) in your knitted HTML** (optional: also include a nicely formatted screeplot)
**A brief summary (nicely formatted and professionally written bulletpoints are fine) highlighting some major takeaways from your PCA that can be gleaned from the biplot (and screeplot, if included)**

```{r setup, include=TRUE, message= FALSE, warning=FALSE}
knitr::opts_chunk$set(echo = TRUE, message = FALSE, warning = FALSE)
library(tidyverse)
library(janitor)
library(here)
 library(ggfortify)
 library(palmerpenguins)
 library(patchwork)
```

```{r}


world_env <- read_csv(here('data', 'world_env_vars.csv')) %>% 
  clean_names() %>% 
  drop_na()

world_env_pca <- world_env %>% 
        #select(-species, -island, -year, -sex) %>% 
        select(accessibility_to_cities:elevation, slope:isothermality, wind, cloudiness) %>% 
        scale() %>% #scales variables based on ranges onto the same spectrum
        prcomp()

world_env_pca
```

```{r}
autoplot(world_env_pca, 
         data = world_env, 
         loadings = TRUE,
         colour = 'country',
         loadings.label = TRUE,
         loadings.colour = 'black',
         loadings.label.colour = 'black',
         loadings.label.vjust = -0.5) + 
        #scale_color_manual(values = c('blue', 'purple', 'orange')) +
        theme_minimal()
```

