---
title: "Task 1 - PCA"
author: "Logan Ossentjuk"
output: 
  html_document:
    code_folding: hide
---

You only need to **use one dataset to perform PCA, create a biplot, and interpret the results.** Whichever dataset you choose, create a finalized HTML (knitted from .Rmd) that includes: 
**A useful descriptive introductory summary (3 - 4 sentences) that helps the audience understand the data (include a citation as necessary) and what you’ll be exploring**
**All of your organized and well-annotated code (with warnings/messages suppressed) you wrote to wrangle data then run PCA, and to create a professional looking PCA biplot that appears (with a caption) in your knitted HTML** (optional: also include a nicely formatted screeplot)
**A brief summary (nicely formatted and professionally written bulletpoints are fine) highlighting some major takeaways from your PCA that can be gleaned from the biplot (and screeplot, if included)**

#Overview: 


```{r setup, include=TRUE, message= FALSE, warning=FALSE}
knitr::opts_chunk$set(echo = TRUE, message = FALSE, warning = FALSE)
library(tidyverse)
library(janitor)
library(here)
library(ggfortify)
library(patchwork)
```

```{r, include= TRUE}

world_env <- read_csv(here('data', 'world_env_vars.csv')) %>% 
  clean_names() %>% 
  drop_na()

world_env_sub <- world_env %>% 
  select(country, accessibility_to_cities, cropland_cover, tree_canopy_cover, rain_mean_annual, rain_wettest_month,temp_annual_range, temp_mean_annual, temp_seasonality, wind, cloudiness) %>% 
  rename(
    "Country" = country,
    "Accessibility To Cities" = accessibility_to_cities,
    "Cropland Cover" = cropland_cover,
    "Tree Canopy Cover" = tree_canopy_cover,
    "Annual Mean Rainfall" = rain_mean_annual,
    "Wettest Month of Rain" = rain_wettest_month, 
    "Annual Temperature Range" = temp_annual_range,
    "Annual Mean Temperature" = temp_mean_annual,
    "Temperature Seasonality" = temp_seasonality,
    "Wind" = wind, 
    "Cloudiness" = cloudiness)


###Running PCA
world_env_pca <- world_env_sub %>% 
  select(-Country) %>% 
  scale() %>% #scales variables based on ranges onto the same spectrum
  prcomp(scale = TRUE)

world_env_pca
```

### PCA Analysis 
```{r}
autoplot(world_env_pca, 
         data = world_env_sub, 
         loadings = TRUE,
         colour = 'darkgreen',
         loadings.label = TRUE,
         loadings.colour = 'goldenrod3',
         loadings.label.colour = 'black',
         loadings.label.vjust = -0.5) + 
        #scale_color_manual(values = c('blue', 'purple', 'orange')) +
        theme_grey()
```

#Summary of Findings: 

 Tree canopy was positively correlated mean and wettest month rainfall, while negatively correlated with temperature seasonality and temperature range. This aligns with our knowledge that trees need a high level of precipitation to thrive. While we cannot infer causality from the correlation, it raises an interesting questions about if the presence of forests serves as a buffer against temperature extremes or if temperature extremes undermine trees' survivability.
- Driest month of rainfall was negatively correlated with temperature range, perhaps pointing to the fact that desert-like environments have large temperature swings throughout the day.
- Rain seasonality was almost completely unrelated to temperature seasonality.
- The variation in accessibility to cities was not captured very well by PCA 1 or PCA 2, indicating a weak relationship between accessibility to cities and the other variables analyzed.

```{r}
